---
title: "HW4"
author: "Eric Winans"
date: "9/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For homework 4, please pull the folder HW4 from the class_notesrepository on GitHub. It contains the pheno.txt* document to be used in this homework. From the pheno.txt file, obtain the following information:
```{r}
pheno <- read.table("pheno.txt", header = TRUE)
```

# 1) How many unique observations are there in the “GENO” variable? (You haven’t used this function in class yet. Try to look for functions.)
```{r}
(unique_geno_obs <- length(unique(pheno$GENO)))
```

# 2) Save as a text file a subset of the pheno data set containing only range 1.
```{r}
write.table(x = pheno[pheno$range == 1,], file = "pheno_range_1.text")
```

# 3) The range (minimum and maximum value) of the “Y”, “M”, and “h4” variables.
```{r}
(range_Y <- range(pheno$Y, na.rm = TRUE))
(range_M <- range(pheno$M, na.rm = TRUE))
(range_h4 <- range(pheno$h4, na.rm = TRUE))
```

# 4) What row contains the maximum value of “h4”?
There are three observations that have the maximum value for "h4" of 5.100. Below I have identified the observations and what "row" in the field they belong to.
```{r}
(max_h4_obs <- which(pheno$h4 == max(pheno$h4, na.rm = TRUE)))
(max_h4_row <- pheno[max_h4_obs,]$row)
```

# 5) Create a loop that prints the mean of the variables “Y”, “M”, “h1”, “h2”, “h3”, and “h4” (it must be done using a loop, e.g., for, while, or repeat).
```{r}
variables <-  pheno[,c("Y", "M", "h1", "h2", "h3", "h4")]
variable_names <- variable.names(variables)
means <- NULL
for (i in variable_names) {
  means[i] <- mean(variables[,i], na.rm = TRUE)
}
means
```

# 6) Create a dataset called pheno2 and replace the values of “Y” that are smaller than 2 by 0.
```{r}
pheno2 <- pheno
pheno2$Y[pheno2$Y<2] <- 0
```

