---
title: "Homework 5"
author: "Eric Winans"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the R example data “mtcars”, do the following (when possible, try using the options available on the tidyverse package):
```{r}
data(mtcars)
str(mtcars)
library(tidyverse)
```


1) Sort the dataset based on the variables “cyl” and “gear” (a single step, cyl is the first variable to be sorted and gear the second one).
```{r}
(mtcars_sorted <- arrange(mtcars, cyl, gear))
```


2) Create a variable called “model” with the names of the rows of this dataset.
```{r}
(mtcars_2 <- mutate(mtcars_sorted, model = row.names(mtcars_sorted)))
```


3) Replace the space of all the models by "_" (e.g., replace “Toyota Corona” by “Toyota_Corona”).
```{r}
(mtcars_3 <- mutate(mtcars_2, model = gsub(" ", "_", mtcars_2$model)))
```


4) Create two datasets, one with mpg_low (by selecting cars with MPG less than 20) and mpg_high (by selecting cars with MGP greater than or equal 20)
```{r}
(mpg_low <- filter(mtcars_3, mpg < 20))
(mpg_high <- filter(mtcars_3, mpg >= 20))
```


5) Using the pipe (%>%) command do all of these steps onto the mtcars dataset: sort by “hp”, remove all columns different than “cyl”, “hp”, " model“, and”gear“, create a new variable named”new_variable" that is the multiplication of “cyl” x “gear”.
```{r}
(mtcars_final <- mtcars_3 %>% 
  arrange(hp) %>% 
  select(cyl, hp, model, gear) %>% 
  mutate(new_variable = cyl * gear))
```

