---
title: "Homework 6"
author: "Eric Winans"
date: "10/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the data flights from the package nycflights13, do the following:
```{r}
library(tidyverse)
library(nycflights13)
data("flights")
head(flights)
```

1- obtain the range, mean, and median of the variable “distance”.
```{r}
range(flights$distance)
mean(flights$distance)
median(flights$distance)
```


2- combine the airlines (also available from nycflights13) with the flights data in a single file.
```{r}
data("airlines")
head(airlines)
airline_flights <- merge(flights, airlines)
head(airline_flights)
```


3- save a file including only American Airlines (Carrier = AA) ordered by arr_delay.
```{r}
airline_flights %>% 
  filter(carrier == "AA") %>% 
  arrange(arr_delay) %>% 
  write.table("AA_flights.txt")
```


4- create boxplots of dep_delay by carrier (one boxplot for each carrier).
```{r}
library(ggplot2)
ggplot(airline_flights, aes(carrier, dep_delay)) +
  geom_boxplot()
```


5- find what is the month with the highest arr_delay. Show this in a graph.
```{r}
#When looking at all values within each month, January (1) contains the highest value for ar_delay
ggplot(airline_flights, aes(month, arr_delay)) +
  geom_point() +
  scale_x_continuous(n.breaks = 12)
```

```{r}
#When plotting the average of arr_delay for each month, July (7) has the highest average arr_delay.
mean_arr_delay <- airline_flights %>% 
  group_by(month) %>% 
  summarise(mean_arr_delay = mean(arr_delay, na.rm = TRUE))

ggplot(mean_arr_delay, aes(month, mean_arr_delay)) +
  geom_col() +
  scale_x_continuous(n.breaks = 12)
```

