---
title: "Homework 7"
author: "Eric Winans"
date: "10/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the pheno.txt inside the HW4 folder, do the following:
```{r}
pheno <- read.table("pheno.txt", header = TRUE)
library(tidyr)
library(tidyverse)
library(ggplot2)
```

1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)
```{r}
#putting all h values into same column
transform_pheno <- pheno %>% 
  pivot_longer(c('h1', 'h2', 'h3', 'h4'), names_to = "measurement", values_to = "value")

#histogram
ggplot(transform_pheno, aes(measurement, value)) + geom_histogram(stat = "identity")
```

2) Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.
```{r}
#putting range and row values into same column
longer_pheno <- transform_pheno %>% 
  pivot_longer(c("range", "row"), names_to = "range_row", values_to = "number")

#boxplot
ggplot(longer_pheno, aes(as.factor(number), Y))+
  geom_boxplot()+
  xlab("values")+
  facet_wrap(vars(range_row), scales = "free_x") 
```

3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error).
```{r}
#averaging Y for each TAXA and filtering by means greater than 12
mean_taxa <- longer_pheno %>% 
  group_by(TAXA) %>% 
  summarise(average = mean(Y)) %>% 
  filter(average > 12)

#barplot with error bars
ggplot(mean_taxa, aes(TAXA, average))+
  geom_col()+
  ylab("mean Y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  geom_errorbar(aes(ymin=average-1, ymax=average+1))
```


The folder HW7 contains one example graph for each question.