---
title: "Lab4"
author: "Eric Winans"
date: "9/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loops and Vectorizatoin

```{r}
vector1 <- runif(100)
vector2 <- runif(100)
matrix1 <- cbind(vector1, vector2)
r1 <- c()
for(i in 1:100){
  r1[i] <- matrix1[i,1] + matrix1[i, 2]
  print(i)
}

matrix2 <- matrix(NA, 100, 10)
matrix3 <- matrix(NA, 1, 10)
for(i in 3:100) {
  for (j in 1:10) {
    matrix2[i, j] <- rnorm(1)
    ex1 <- rnorm(10)
    matrix3 <- rbind(matrix3, rnorm(10))
  }
}
```

__The apply family__:
[s,l,m,r,t,v]apply: nothing but a loop internally

```{r}
#apply(X, MARGIN, FUN, ...)
r2 <- apply(matrix1, 1, sum)
body(apply)
```

```{r}
#apply(FUN, ..., MoreArgs = NULL, SIMPLIFY = TRUE, USE.NAMES = TRUE)
r3 <- mapply(sum, vector1, vector2)
```

```{r}
#lapply(X, FUN, ...)
lapply(as.list(vector1), function(a){ paste("printing number:", a)})
#sapply(X, FUN, ..., simplify = TRUE, USE.NAMES = TRUE)
sapply(vector1, function(x){ paste("printing number:", x)})
```

```{r}
data("iris")
#tapply(X, INDEX, FUN = NULL, ..., default = NA, simplify = TRUE)
tapply(iris$Sepal.Length, iris$Species, mean)
```

# Vectorization

```{r}
r4 <- vector1 + vector2

all(r1==r2 & r3==r4)
all.equal(r1, r2, r3, r3)
```


# Profiling R code

Make our code to run faster, but only after having it well designed!

True: for example, preallocating memory can speed things up
Fals: apply functions are inherently faster than for loops

Suppose you make a loop run 5 times faster. That sounds like a huge improvement, but if that loop only takes 10% or the total time, it's stil only a 8% speedup overall. To make slow code faster, we need accurate information about what is making our code slow. 

```{r}
my_for <- function(matrix1) {
  for (i in 1:100) {
    r1[i] <- matrix1[i, 1] + matrix1[i, 2]
  }
  return(r1)
}

my_apply <- function(matrix1) {
  return(apply(matrix1, 1, sum))
}

system.time(my_for(matrix1))

system.time(my_apply(matrix1))

system.time(r4 <- vector1 + vector2)
```

```{r}
install.packages("microbenchmark")
library(microbenchmark)
microbenchmark::microbenchmark(my_for(matrix1), 
                               my_apply(matrix1),
                               {r4 <- vector1 + vector2},
                               times = 100)
```


```{r}
install.packages("profvis")
library(profvis)
profvis({
  data(diamonds, package = "ggplot2")
  plot(price ~ carat, data = diamonds)
  m <- lm(price ~ carat, ddata = diamonds)
  abline(m, col = "red")
})

times <- 4e5
cols <- 150
data <- 
  as.data.frame(x = matrix(rnorm(times * cols, mean = 5),
                           ncol = cols))
data <- cbind(id = paste0("g", seq_len(times)), data)

profvis({
  #store in another variable for this run
  data1 <- data
  
  #get colum means
  means <- apply(data1, margin, ...)
  # Subtract mean from each column
  for (i in seq_along(means)) {
    data1[, names(data1) != "id"][, i] <- 
      data1[, names(data1) != "id"][, i] - means[i]
  }
})
```

