---
title: "Functions"
author: "Eric Winans"
date: "9/11/2020"
output: html_document
---

# What is a function?
A function is a set of organized statements used to conduct a specific task. R has a large number of built-in functions and the user can create their own functions.

*  __Function Name__

*  __Function Body__ - The function body contains a collection of statement that defines what the function does

*  __Return Value__ - The return value of a function is the last expression in the function body to be evaluated


```{r}
temp_F <- 72
base_temp <- 32

fahrenhite_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - base_temp) * 5 / 9
  return(temp_C)
}
```

# Environment

```{r eval=FALSE}
# Global Environment
 temp_F
# Scope of the function
 temp_C
 
 mean <- function(arg){
   x <- sum(arg)
   return(x)
 }
```

# Lazy Evaluation of Function
```{r eval=FALSE}
new_function <- function(a, b) {
  print(a^2)
  print(a)
  princomp(b)
}
# Evaluate the function without supplying one of the arguments
new_function(b = 6, a = 1000)
```

# create a function with default values
```{r}
display <- function(a = 1, b = 20, c = 3){
  result <- c(a, b, c)
  names(result) <- c("a", "b", "c")
  return(result)
}

# no arguments
display()

input_1 <- 20
mySum <- function(input_1, input_2 = 10) {
  output <- input_1 + input_2
  return(output)
}
mySum(input_1 = 1)
```

```{r}
# Anonymous function syntax
(function(x) x *10)(10)
```

```{r}
pow <- function(x, y) {
  # function to print x raised to the power y
  result <- x^y
  print(paste(x,"raised to the power", y, "is", result))
}
pow(x = 2, y = 5)
```


#Implementing math in R

$$\mathrm{Var}(x) = \frac{1}{n - 1} \sum_{i=1}^n (x_i - \bar{x}) ^2 \text{,}$$


```{r}
a <- 1:10
var(a)

(1 / 9) * ((a - mean(a)) ^2)

b <- 20:30

my_var <- function(vector){
  r1 <- 1 / (length(vector) - 1)
  r2 <- (vector - mean(vector))^2
  r1 * sum(r2)
  final <- r1 * sum(r2)
  
  x <- 10
  return(list(final = final, x = x))
}
my_var(a)
my_var(b)
```



